Efficiently Handling Guards in the low level design of RPython's tracing JIT

### 2012-07-20 talk/vmil2012/paper.tex
more low level

Code generation consists of two passes over the lists of instructions, a
backwards pass to calculate live ranges of IR-level variables and a forward one
to emit the instructions. During the forward pass IR-level variables are
assigned to registers and stack locations by the register allocator according
to the requirements of the to be emitted instructions. Eviction/spilling is
performed based on the live range information collected in the first pass. Each
IR instruction is transformed into one or more machine level instructions that
implement the required semantics, operations withouth side effects whose result
is not used are not emitted. Guards instructions are transformed into fast
checks at the machine code level that verify the corresponding condition.  In
cases the value being checked by the guard is not used anywhere else the guard
and the operation producing the value can merged, reducing even more the
overhead of the guard. \bivab{example for this}

Each guard in the IR has attached to it a list of the IR-variables required to
rebuild the execution state in case the trace is left through the side-exit
corresponding to the guard. When a guard is compiled, additionally to the
condition check two things are generated/compiled. First a special
data structure is created that encodes the information provided by the register
allocator about where the values corresponding to each IR-variable required by
the guard will be stored when execution reaches the code emitted for the
corresponding guard. \bivab{go into more detail here?!} This encoding needs to
be as compact as possible to maintain an acceptable memory profile.

\bivab{example goes here}

Second a trampoline method stub is generated. Guards are usually implemented as
a conditional jump to this stub, jumping to in case the guard condition check
does not hold and a side-exit should be taken. This stub loads the pointer to
the encoding of the locations mentioned above, preserves the execution state
(stack and registers) and then jumps to generic bail-out handler that is used
to leave the compiled trace if case of a guard failure.

>>>>

After optimization the resulting trace is handed to the backend to be compiled
to machine code. The compilation phase consists of two passes over the lists of
instructions, a backwards pass to calculate live ranges of IR-level variables
and a forward one to emit the instructions. During the forward pass IR-level
variables are assigned to registers and stack locations by the register
allocator according to the requirements of the to be emitted instructions.
Eviction/spilling is performed based on the live range information collected in
the first pass. Each IR instruction is transformed into one or more machine
level instructions that implement the required semantics, operations withouth
side effects whose result is not used are not emitted. Guards instructions are
transformed into fast checks at the machine code level that verify the
corresponding condition.  In cases the value being checked by the guard is not
used anywhere else the guard and the operation producing the value can merged,
reducing even more the overhead of the guard. \bivab{example for this}

Each guard in the IR has attached to it a list of the IR-variables required to
rebuild the execution state in case the trace is left through the side-exit
corresponding to the guard. When a guard is compiled, additionally to the
condition check two things are generated/compiled. First a special data
structure called \emph{low-level resume data} is created that encodes the
information provided by the register allocator about where the values
corresponding to each IR-variable required by the guard will be stored when
execution reaches the code emitted for the corresponding guard. \bivab{go into
more detail here?!} This encoding needs to be as compact as possible to
maintain an acceptable memory profile.

\bivab{example for low-level resume data goes here}

Second a piece of code is generated for each guard that acts as a trampoline.
Guards are implemented as a conditional jump to this trampoline. In case the
condition checked in the guard fails execution and a side-exit should be taken
execution jumps the the trampoline. In the trampoline the pointer to the
\emph{low-level resume data} is loaded and jumps to generic bail-out handler
that is used to leave the compiled trace if case of a guard failure.
